---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0.1
#ESSENTIALS
title: '**Leveraging allelic series to enhance QTL detection**'
author:
  - name: '**Frederick J. Boehm**'
    affil: 1
    main: true
    orcid: '0000-0002-1644-5931'
    email: frederick.boehm@umassmed.edu
  - name: Clare M. Smith
    affil: 2
  - name: Megan Proulx
    affil: 1
  - name: Daniel M. Gatti
    affil: 3
  - name: Michael Kiritsy
    affil: 1
  - name: Sherry Kurtz
    affil: 4
  - name: Karen Elkins
    affil: 4
  - name: Gillian Beamer
    affil: 5
  - name: Christopher M. Sassetti
    affil: 1
affiliation:
  - num: 1
    address: Microbiology and Physiological Systems, University of Massachusetts Medical School
  - num: 2
    address: Duke University
  - num: 3
    address: College of the Atlantic
  - num: 4
    address: Food and Drug Administration
  - num: 5
    address: Tufts University
main_findings:
  - "Inferring QTL allelic series reduces the phenotype modeling parameter space"
logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logoright_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logocenter_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/qr-code-black.png
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: ../research.bib
link-citations: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

# Introduction

- In CC and DO, we typically work with 8 founder allele probabilities

$${Trait} = p_{A}b_A +  
p_{B}b_B +
p_{C}b_C +
p_{D}b_D +
p_{E}b_E +
p_{F}b_F +
p_{G}b_G +
p_{H}b_H + 
\epsilon$$

- Standard QTL methods allow every founder to have its own allele
- However, founders often share alleles
- Thus, in DO and CC, some QTL have fewer than 8 alleles
- Using this in our QTL scans may enhance QTL detection


```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

## Objectives

1. Infer QTL allelic series in CC mice
1. Use allelic series to map QTL in DO mice


# Methods

- We measured *M. tuberculosis* burden in lungs and spleen in a DO cohort (___ mice) and a CC cohort (53 lines)

- We mapped QTL in CC cohort with standard methods

- We used Tree-based Inference of Multi-allelism with Bayesian Regression (TIMBR) [@crouse2018timbr] to infer allelic series at CC QTL

- For each CC QTL, we scanned the DO cohort with reduced models that reflected the CC QTL allelic series

    - Example of reducing a model with inferred allelic series
    For the allelic series: 0,1,1,0,2,1,0,0:
    1. Set $p_0 = p_A + p_D + p_G + p_H$,
    $p_1 = p_B + p_C + p_F$
    $p_2 = p_E$
    2. Use model $$Y = \beta_0p_0 + \beta_1 p_1 + \beta_2p_2 + \epsilon$$ in QTL scan




**_Now on to the results!_**

# Results




# References
