---
title: "Enhancing QTL mapping with allelic series inferences"
author: "Fred Boehm"
institute: "U. Mass. Medical School"
date: "June 24, 2020 (updated: `r Sys.Date()`)"
bibliography: ../research.bib
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true  
      countIncrementalSlides: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r setup}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
options(future.globals.maxSize= 8 * 1024^3)
```


```{r}
PATH_TO_DDATA <- "~/work/research/qtl2tb/analysis/data/derived_data"
load(file.path(PATH_TO_DDATA, "reduced_map_qtl2_mapping_objects.Rdata"))
library(magrittr)
load(file.path(PATH_TO_DDATA, "fda_qtl_input.Rdata"))
do_mapdf <- qtl2convert::map_list_to_df(map) %>%
  dplyr::rename(pos_Mb = pos)
cc_mapdf <- qtl2convert::map_list_to_df(map.1) %>%
  dplyr::rename(pos_cM = pos) %>%
  dplyr::left_join(do_mapdf) 
```

```{r, scan1-cc}
cc_kinship <- qtl2::calc_kinship(genoprobs.1, "loco")
cc_pre_pheno <- read.csv(file.path(PATH_TO_DDATA, "pheno_all_clinical_v4_suggestive.csv"))
cc_pheno<- as.matrix(cc_pre_pheno[, -1])
rownames(cc_pheno) <- cc_pre_pheno[ , 1]

cc_s1 <- qtl2::scan1(genoprobs = genoprobs.1, 
                     pheno = cc_pheno, 
                     kinship = cc_kinship, 
                     addcovar = addcovar, 
                     reml = TRUE
                     )
cc_peaks <- qtl2::find_peaks(cc_s1, map = qtl2convert::map_df_to_list(cc_mapdf, pos_column = "pos_Mb"), threshold = 5, prob = 0.95) %>%
  tibble::as_tibble() %>%
  dplyr::rename(cc_peak_bayes_lower = ci_lo, cc_peak_bayes_upper = ci_hi)
```

```{r}
# permutations
dfn <- file.path(PATH_TO_DDATA, "cc_s1perm_out.rds")
if (!file.exists(dfn)){
  s1perm_out <- qtl2::scan1perm(genoprobs = genoprobs.1, pheno = cc_pheno, kinship = cc_kinship, addcovar = addcovar, reml = TRUE, n_perm = 1000, cores = parallel::detectCores())
  saveRDS(s1perm_out, file = dfn)  
} else {
  s1perm_out <- readRDS(dfn)
}
```

```{r load-genotypes}
fn36 <- file.path(PATH_TO_DDATA, "cc36.rds")
gp <- readRDS(fn36)
```

```{r TIMBR-setup}
future::plan("multiprocess")
fn <- file.path(PATH_TO_DDATA, "cc-timbr-clinical-M1.rds")
if (!file.exists(fn)){
  prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
    prior.alpha.type = "gamma",
    prior.alpha.shape = 1,
    prior.alpha.rate = 2.333415)
  prior_M2 <- list(model.type = "crp",
                  prior.alpha.type = "gamma",
                  prior.alpha.shape = 2.3009322,
                  prior.alpha.rate = 0.7488104
                  )
  library(TIMBR)
  data(mcv.data) # get A matrix
  set.seed(3411192) # to ensure getting the same samples with TIMBR
  t_out <- cc_peaks %>%
    dplyr::mutate(marker_index = purrr::map2_int(.x = pos, .y = chr, .f = function(x, y){
      ppcm <- cc_mapdf %>%
        dplyr::filter(chr == y, x == pos_Mb) %>%
        dplyr::select(pos_cM) %>%
        unlist()
      
      return(which(map.1[[y]] == ppcm))
      })) %>%
    dplyr::filter(chr != "X") %>% # No TIMBR for X chr yet!
    dplyr::mutate(timbr = furrr::future_pmap(.l = list(lodcolumn, chr, marker_index), .f = function(lodcolumn, chr, marker_index){
      phe_pre <- cc_pheno[ , colnames(cc_pheno) == lodcolumn, drop = FALSE]
      phe <- phe_pre[!is.na(phe_pre), , drop = FALSE]
      pr_pre <- gp[[chr]][, , marker_index]
      pr <- pr_pre[rownames(pr_pre) %in% rownames(phe), ]
      cov_pre <- cbind(1, addcovar)
      cov_pre2 <- cov_pre[rownames(cov_pre) %in% rownames(phe), ]
      cov <- qtl2:::drop_depcols(cov_pre2)
      TIMBR::TIMBR(y = phe, 
                   prior.D = list(P = pr, 
                                  A = mcv.data$prior.D$A, # works for autosomes
                                  fixed.diplo = FALSE), 
                   prior.M = prior_M1, 
                   #samples = 3, samples.ml = 3,
                   samples = 10000, samples.ml = 10000,
                   Z = cov, 
                   verbose = TRUE
                   )  
  }))
  saveRDS(t_out, fn)
} else {
  t_out <- readRDS(fn)
}

```

```{r}
(t2 <- t_out %>%
  dplyr::mutate(most_probable_series = purrr::map_chr(.x = timbr, 
                                                      .f = function(x)names(x$p.M.given.y[1])), 
                most_probable_series_prob = purrr::map_dbl(.x = timbr, 
                                                           .f = function(x)x$p.M.given.y[1]),
                second_most_probable_series = purrr::map_chr(.x = timbr, 
                                                      .f = function(x)names(x$p.M.given.y[2])), 
                second_most_probable_series_prob = purrr::map_dbl(.x = timbr, 
                                                           .f = function(x)x$p.M.given.y[2])) %>%
  dplyr::mutate(top_pairs = purrr::map(.x = timbr, .f = qtl2tb::find_top_pairs))
)
```

```{r}
afn <-  file.path(PATH_TO_DDATA, "cc-timbr-clinical-M1-with-reduced-peaks.rds")
if (!file.exists(afn)){
  t3 <- t2 %>%
  dplyr::mutate(peaks_reduced_s1out = furrr::future_map2(.x = most_probable_series, 
                                                         .y = second_most_probable_series,
                                                        .f = function(x, y){
    if (x != "0,0,0,0,0,0,0,0"){
      as <- qtl2tb::convert_character_series_to_binary_series(x)
          } else {
      as <- qtl2tb::convert_character_series_to_binary_series(y)
          }
      pr <- reducedscan::reduce_probs(probs, allelic_series = as)
      out <- qtl2::scan1(genoprobs = pr, 
                       pheno = pheno_rz, 
                       kinship = K, 
                       addcovar = covar, 
                       reml = TRUE
                       )
    
    return(out)
  })
  
                  )

saveRDS(t3, afn)
}
t3 <- tibble::as_tibble(readRDS(afn))
```



# Mtb-related QTL in CC

- 52 CC lines
- Measured 11 "clinical" traits
  - spleen CFU, lung CFU, spleen CEQ
  - 8 cytokines


???

Broad peaks and limited detection power

---

# Mtb-related QTL in CC

```{r}
cc_peaks %>%
  dplyr::mutate(pvalue = purrr::map2_dbl(.x = lodcolumn, .y = lod, .f = function(x, y){
    maxlods <- s1perm_out[, colnames(s1perm_out) == x]
    mean(maxlods >= y)
  })) %>%
  dplyr::filter(pvalue < 0.37) %>%
  dplyr::select(- lodindex, - cc_peak_bayes_lower, - cc_peak_bayes_upper) %>%
  dplyr::rename(cc_trait = lodcolumn) %>%
  gt::gt() %>%
    gt::fmt_number(
    columns = dplyr::vars(lod, pos),
    decimals = 1
  ) %>%
  gt::tab_style(style = list(
      gt::cell_fill(color = "lightcyan"),
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = dplyr::vars(cc_trait, chr, pos, lod, pvalue),
      rows = (cc_trait == "VEGF" & chr == 4))) %>%
  gt::tab_style(style = list(
      gt::cell_fill(color = "pink"),
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = dplyr::vars(cc_trait, chr, pos, lod, pvalue),
      rows = (pvalue < 0.05)))
  

```


---

# Mtb-related QTL in DO

- 763 mice 
- 5 traits: lung CFU, spleen CFU, percent normal lung, peak body weight, euthanization body weight

---

# Mtb-related QTL in DO

```{r, scan1}
s1out <- qtl2::scan1(genoprobs = probs, 
                     pheno = pheno_rz, 
                     kinship = K, 
                     addcovar = covar, 
                     reml = TRUE, cores = 3
                     )
do_peaks <- qtl2::find_peaks(s1out, map = map, threshold = 5)
do_s1perm_out <- readRDS(file = file.path(PATH_TO_DDATA, "do_s1perm_out.rds"))
```

```{r}
do_peaks %>%
  dplyr::mutate(pvalue = purrr::map2_dbl(.x = lodcolumn, .y = lod, .f = function(x, y){
    maxlods <- do_s1perm_out[, colnames(s1perm_out) == x]
    mean(maxlods >= y)
  })) %>%
  dplyr::filter(pvalue < 0.37)
  dplyr::select(- lodindex, - cc_peak_bayes_lower, - cc_peak_bayes_upper) %>%
  dplyr::rename(cc_trait = lodcolumn) %>%
  gt::gt() %>%
    gt::fmt_number(
    columns = dplyr::vars(lod, pos),
    decimals = 1
  ) %>%
  gt::tab_style(style = list(
      gt::cell_fill(color = "lightcyan"),
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = dplyr::vars(cc_trait, chr, pos, lod, pvalue),
      rows = (cc_trait == "VEGF" & chr == 4))) %>%
  gt::tab_style(style = list(
      gt::cell_fill(color = "pink"),
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = dplyr::vars(cc_trait, chr, pos, lod, pvalue),
      rows = (pvalue < 0.05)))
  
```





---

# What are allele effects? 

- In CC and DO, we work with 8 founder allele probabilities


$${Trait} = p_{A}b_A +  
p_{B}b_B +
p_{C}b_C +
p_{D}b_D +
p_{E}b_E +
p_{F}b_F +
p_{G}b_G +
p_{H}b_H + 
\epsilon$$

---

# What are allele effects?



???

- Standard QTL mapping methods allow for every founder to have its own allele
- However, founders often share (functional) alleles
- Thus, in DO and CC, some QTL have fewer than 8 distinct alleles
- QTL exhibit contrasts in allele effects
  - "Contrast" is a pairwise difference in mean value
- How many alleles are present here? 


---

# What are allele effects?




---

# TIMBR Methods for inferring allelic series in QTL studies [@crouse2020inferring]




---

# Cis DO peaks


```{r}
bfn <-  file.path(PATH_TO_DDATA, "cc-timbr-clinical-M1-with-reduced-peaks-cis.rds")
if (!file.exists(bfn)){
  t4 <- t3 %>%
  dplyr::mutate(peaks_reduced = purrr::map(.x = peaks_reduced_s1out, 
                                           .f = function(x) {
                                             if (!is.null(dim(x))){
                                               out <- qtl2::find_peaks(x, 
                                                                       map = map,
                                                                       threshold = 1, 
                                                                       prob = 0.95) # might change threshold value
                                               } else {out <- NA}
                                             return(out)
                                           })) %>%
  dplyr::mutate(cis_peaks = purrr::pmap(.l = list(peaks_reduced, chr, cc_peak_bayes_lower, cc_peak_bayes_upper), 
                                        .f = function(x, y, z, w){
                                              if (!is.null(dim(x))){
                                                out <- x %>%
                                                  dplyr::filter(chr == y, 
                                                                pos < w, 
                                                                pos > z
                                                                )
                                                } else {NA -> out}
                                              return(out)
                                              }
                                        )
                )
saveRDS(t4, bfn)

} else {
  t4 <- readRDS(bfn)
}
```

```{r}
t4 %>%
  dplyr::mutate(cis2 = purrr::pmap(.l = list(cis_peaks, lodcolumn, pos), .f = function(x, y, z){
    x %>%
      dplyr::mutate(cc_trait = y, cc_trait_pos = z)
  })) %>%
  dplyr::select(cis2) %>%
  purrr::map_df(dplyr::bind_rows) %>%
  dplyr::select(- ci_lo, - ci_hi, - lodindex) %>%
  dplyr::filter(lod > 3
                ) %>%
  dplyr::rename(do_trait = lodcolumn) %>%
  gt::gt() %>%
   gt::fmt_number(
    columns = dplyr::vars(pos, lod, cc_trait_pos),
    decimals = 1
  ) %>%
  gt::tab_style(style = list(
      gt::cell_fill(color = "lightcyan"),
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = dplyr::vars(do_trait, chr, pos, lod, cc_trait, cc_trait_pos),
      rows = (do_trait == "spleen_mtb")))
```


---

# xxx


---
class: inverse, center, middle

# References

